---
title: "03_ordination_diversity"
author: "Daniel"
date: "27/09/2022"
output: github_document
---


```{r message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls()) #I always start scritps this way to make sure I have a clean R environment
library('tidyr')
library('dplyr')
library('metacoder')
library('ggplot2')
library('agricolae')
library('vegan')
library('MicrobiotaProcess')
library('phyloseq')

obj <- readRDS('data/taxmap_object.rds') #loads the taxmap object created in script 02_metacoder_heat_trees
sample <- read.csv('data/location_ctrl.csv')
```


```{r simpleBoxplotANOVAbetweenSeason}
sample$inv_simp <- diversity(obj$data$tax_data[, sample$sampleID],
                             index = "invsimpson",
                             MARGIN = 2) # What orietation the matrix is in
ggplot(sample, aes(x = SEASON, y = inv_simp)) +
  geom_boxplot()
anova_result <- aov(inv_simp ~ SEASON, sample)
summary(anova_result)
tukey_result <- HSD.test(anova_result, "SEASON", group = TRUE)
print(tukey_result)

group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
ggplot(sample, aes(x = SEASON, y = inv_simp)) +
  geom_text(data = data.frame(),
            aes(x = rownames(group_data), y = max(sample$inv_simp) + 1, label = group_data$groups),
            col = 'black',
            size = 10) +
  geom_boxplot() +
  ggtitle("Inverse Simpson diversity") +
  xlab("Season") +
  ylab("Inverse simpson index")
```

We can also see differences between sampling days.

```{r simpleBoxplotANOVAbetweenDays}
ggplot(sample, aes(x = day, y = inv_simp)) +
  geom_boxplot()
anova_result <- aov(inv_simp ~ day, sample)
summary(anova_result)
tukey_result_day <- HSD.test(anova_result, "day", group = TRUE)
print(tukey_result_day)

group_data_day <- tukey_result_day$groups[order(rownames(tukey_result_day$groups)),]
ggplot(sample, aes(x = day, y = inv_simp)) +
  geom_text(data = data.frame(),
            aes(x = rownames(group_data_day), y = max(sample$inv_simp) + 1, label = group_data_day$groups),
            col = 'black',
            size = 10) +
  geom_boxplot() +
  ggtitle("Inverse Simpson diversity") +
  xlab("Day") +
  ylab("Inverse simpson index")
```

No difference between sampling days.
We can also calculate several diversity metrics for sampling season and day using phyloseq but we must convert the taxmap object 'obj' to phyloseq 
```{r multipleDiversityIndices}
ps_obj <- metacoder::as_phyloseq(obj,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID")

#normally, I load every package I need at the start of the script, but this may cause function masking problems (e.g. both MicrobiotaProcess and Metacoder has a 'as_phyloseq' function and you have to specify to R which package you want to use -  see above: metacoder::as_phyloseq - this means I am telling R to use the 'as_phyloseq' function from the metacoder package)
plot_richness(ps_obj, color = "SEASON", x = "site") #phyloseq function
plot_richness(ps_obj, color = "day", x = "SEASON")

```
Figures above OK but we can do better.
I am using the MicrobiotaProcess next

```{r MBPPlotsDiversityIndices}
alphaobj <- get_alphaindex(ps_obj)
p_alpha <- ggbox(alphaobj, geom="violin", factorNames="SEASON", indexNames = c('Observe', 'Shannon', 'Simpson'),  signifmap = TRUE) +
  scale_fill_manual(values=c("orange", "darkcyan"))+
  theme(strip.background = element_rect(colour=NA, fill="grey"))
p_alpha
```
And look at rarefaction curves for both seasons:
```{r MBPRarefactionCurvesPerSeason}
alphaobj <- get_alphaindex(ps_obj)
head(as.data.frame(alphaobj))

rareres <- get_rarecurve(obj=ps_obj, chunks=400)

prare2 <- ggrarecurve(obj=rareres,
                      factorNames="SEASON",
                      shadow=FALSE,
                      indexNames="Observe"
) +
  scale_color_manual(values=c("orange", "darkcyan"))+
  theme_bw()+
  theme(axis.text=element_text(size=8), panel.grid=element_blank(),
        strip.background = element_rect(colour=NA,fill="grey"),
        strip.text.x = element_text(face="bold"))

prare2
```

We can also do a Principal Coordinate Analysis to see not onlyl the differences between seasons, but what BINs are the main drivers of these differences

```{r MBPPCoAPlots}
# distmethod
# "unifrac",  "wunifrac", "manhattan", "euclidean", "canberra", "bray", "kulczynski" ...(vegdist, dist)
pcoares <- get_pcoa(obj=ps_obj, distmethod="bray", method="hellinger")
# Visualizing the result
pcoaplot1 <- ggordpoint(obj=pcoares, biplot=TRUE, speciesannot=TRUE,
                        factorNames=c("SEASON"), ellipse=TRUE) +
  scale_color_manual(values=c("orange", "darkcyan")) +
  scale_fill_manual(values=c("orange", "darkcyan"))
# first and third principal co-ordinates
pcoaplot2 <- ggordpoint(obj=pcoares, pc=c(1, 3), biplot=TRUE, speciesannot=TRUE,
                        factorNames=c("SEASON"), ellipse=TRUE) +
  scale_color_manual(values=c("orange", "darkcyan")) +
  scale_fill_manual(values=c("orange", "darkcyan"))
pcoaplot1 | pcoaplot2
```

```{r MBPTaxonAbundanceTOP10}
classtaxa <- get_taxadf(obj=ps_obj, taxlevel=3)
# The 10 most abundant taxonomy will be visualized by default (parameter `topn=10`). 
pclass <- ggbartax(obj=classtaxa, facetNames="SEASON", topn=10) +
  xlab(NULL) +
  ylab("relative abundance (%)") +
  scale_fill_manual(values=c(colorRampPalette(RColorBrewer::brewer.pal(12,"Set3"))(31))) +
  guides(fill= guide_legend(keywidth = 0.5, keyheight = 0.5))
pclass
```

```{r MBPTaxonAbundanceTOP15}
#note the flag "count=TRUE", this shows now total reads, rather than proportion %
pclass2 <- ggbartax(obj=classtaxa, count=TRUE, facetNames="SEASON", topn=15) +
  xlab(NULL) +
  ylab("count reads") +
  scale_fill_manual(values=c(colorRampPalette(RColorBrewer::brewer.pal(12,"Set3"))(31))) +
  guides(fill= guide_legend(keywidth = 0.5, keyheight = 0.5))
pclass2
```

These following chunks might be best to compare metabarcoding vs traditional data but its OK to visualize below.
Way too many species difference between both sampling strategies and so the plots are messy  - after Kruskal, Wilcox and LDA, 264 taxa are discriminative between wet and dry seasons.

```{r}
library('coin')#for kruskal and wilcox test
# Since the effect size was calculated by randomly re-sampling, 
# the seed should be set for reproducibly results.
set.seed(1024)
deres <- diff_analysis(obj = ps_obj, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
deres
```

```{r MessyFigure}
diffclade_p <- ggdiffclade(
  obj=deres, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=7, #taxonomy level from 1 to 7 phylum:class:order:family:subfamily:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("orange", "darkcyan")
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_p
```

```{r}
ps_obj %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^c_", f)) %>% ggdiffbox(colorlist=c("orange", "darkcyan"), notch = FALSE)
```


```{r}
diffbox <- ggdiffbox(obj=deres, box_notch=FALSE, 
                     colorlist=c("orange", "darkcyan"), l_xlabtext="relative abundance", ytextsize = 0.5)
diffbox
```

```{r}
es_p <- ggeffectsize(obj=deres, 
                     lineheight=0.1,
                     linewidth=0.3) + 
  scale_color_manual(values=c("orange","darkcyan")) 

es_p
```