---
title: "03_ordination_diversity"
author: "Daniel"
date: "27/09/2022"
output: github_document
#editor_options: 
  #chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls()) #I always start scritps this way to make sure I have a clean R environment
library('tidyr')
library('dplyr')
library('metacoder')
library('ggplot2')
library('agricolae')
library('vegan')
library('MicrobiotaProcess')
library('phyloseq')
library('ggtree')

obj <- readRDS('data/taxmap_object.rds') #loads the taxmap object created in script 02_metacoder_heat_trees

obj %>%  metacoder::filter_taxa(taxon_names %in% c("Lepidoptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> leps #we will create separate files for each order as this simplifies downstream analysis in microbiotaprocess - until I find a way to filter taxa on the mpse object directly.

obj %>%  metacoder::filter_taxa(taxon_names %in% c("Coleoptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> coleo
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Diptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> dips
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Hymenoptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> bees #i know its hymenoptera
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Hemiptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> hemi
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Blattodea"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> blats 

sample <- read.csv('data/location_ctrl.csv')
```

The chunk above now gave us a filtered taxmap object for each of the ForestGEO focal groups (plus Diptera because I think its interesting).
Downstream analysis will make use of these filtered data sets, but we continue with general info (whole data set) between seasons.

```{r simpleBoxplotANOVAbetweenSeason}
sample$inv_simp <- diversity(obj$data$tax_data[, sample$sampleID],
                             index = "invsimpson",
                             MARGIN = 2) # What orietation the matrix is in
ggplot(sample, aes(x = SEASON, y = inv_simp)) +
  geom_boxplot()
anova_result <- aov(inv_simp ~ SEASON, sample)
summary(anova_result)
tukey_result <- HSD.test(anova_result, "SEASON", group = TRUE)
print(tukey_result)

group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
ggplot(sample, aes(x = SEASON, y = inv_simp)) +
  geom_text(data = data.frame(),
            aes(x = rownames(group_data), y = max(sample$inv_simp) + 1, label = group_data$groups),
            col = 'black',
            size = 10) +
  geom_boxplot() +
  ggtitle("Inverse Simpson diversity") +
  xlab("Season") +
  ylab("Inverse simpson index")
```

We can also see differences between sampling days although there is none. 

```{r simpleBoxplotANOVAbetweenDays}
ggplot(sample, aes(x = day, y = inv_simp)) +
  geom_boxplot()
anova_result <- aov(inv_simp ~ day, sample)
summary(anova_result)
tukey_result_day <- HSD.test(anova_result, "day", group = TRUE)
print(tukey_result_day)

group_data_day <- tukey_result_day$groups[order(rownames(tukey_result_day$groups)),]
ggplot(sample, aes(x = day, y = inv_simp)) +
  geom_text(data = data.frame(),
            aes(x = rownames(group_data_day), y = max(sample$inv_simp) + 1, label = group_data_day$groups),
            col = 'black',
            size = 10) +
  geom_boxplot() +
  ggtitle("Inverse Simpson diversity") +
  xlab("Day") +
  ylab("Inverse simpson index")
```

We can also calculate several diversity metrics for sampling season and day using phyloseq but we must convert the taxmap object 'obj' to phyloseq 

```{r multipleDiversityIndices}
ps_obj <- metacoder::as_phyloseq(obj,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID")

#normally, I load every package I need at the start of the script, but this may cause function masking problems (e.g. both MicrobiotaProcess and Metacoder have a 'as_phyloseq' function and you have to specify to R which package you want to use -  see above: metacoder::as_phyloseq - this means I am telling R to use the 'as_phyloseq' function from the metacoder package)

plot_richness(ps_obj, color = "SEASON", x = "site") #phyloseq function
plot_richness(ps_obj, color = "day", x = "SEASON")

```

Figures above OK but we can do better.
I am using the MicrobiotaProcess next

```{r MBPPlotsDiversityIndices}
alphaobj <- get_alphaindex(ps_obj)
p_alpha <- ggbox(alphaobj, geom="violin", factorNames="SEASON", indexNames = c('Observe', 'Shannon', 'Simpson'),  signifmap = TRUE) +
  scale_fill_manual(values=c("goldenrod", "steelblue"))+
  theme(strip.background = element_rect(colour=NA, fill="grey"))
p_alpha
```
And look at rarefaction curves for both seasons:
```{r MBPRarefactionCurvesPerSeason}
alphaobj <- get_alphaindex(ps_obj)
head(as.data.frame(alphaobj))

rareres <- get_rarecurve(obj=ps_obj, chunks=400)

prare2 <- ggrarecurve(obj=rareres,
                      factorNames="SEASON",
                      shadow=FALSE,
                      indexNames="Observe"
) +
  scale_color_manual(values=c("goldenrod", "steelblue"))+
  theme_bw()+
  theme(axis.text=element_text(size=8), panel.grid=element_blank(),
        strip.background = element_rect(colour=NA,fill="grey"),
        strip.text.x = element_text(face="bold"))

prare2
```

We can also do a Principal Coordinate Analysis to see not only the differences between seasons, but what BINs are the main drivers of these differences

```{r MBPPCoAPlots}
# distmethod
# "unifrac",  "wunifrac", "manhattan", "euclidean", "canberra", "bray", "kulczynski" ...(vegdist, dist)
pcoares <- get_pcoa(obj=ps_obj, distmethod="bray", method="hellinger")
# Visualizing the result
pcoaplot1 <- ggordpoint(obj=pcoares, biplot=TRUE, speciesannot=TRUE,
                        factorNames=c("SEASON"), ellipse=TRUE) +
  scale_color_manual(values=c("goldenrod", "steelblue")) +
  scale_fill_manual(values=c("goldenrod", "steelblue"))
# first and third principal co-ordinates
pcoaplot2 <- ggordpoint(obj=pcoares, pc=c(1, 3), biplot=TRUE, speciesannot=TRUE,
                        factorNames=c("SEASON"), ellipse=TRUE) +
  scale_color_manual(values=c("goldenrod", "steelblue")) +
  scale_fill_manual(values=c("goldenrod", "steelblue"))

pcoaplot1 

pcoaplot2

pcoaplot1 | pcoaplot2
```

```{r MBPTaxonAbundanceTOP10}


classtaxa <- get_taxadf(obj=ps_obj, taxlevel=4)
# The 10 most abundant taxonomy will be visualized by default (parameter `topn=10`). 
pclass <- ggbartax(obj=classtaxa, facetNames="SEASON", topn=10) +
  xlab(NULL) +
  ylab("relative abundance (%)") +
  scale_fill_manual(values=c(colorRampPalette(RColorBrewer::brewer.pal(12,"Set3"))(31))) +
  guides(fill= guide_legend(keywidth = 0.5, keyheight = 0.5))

#note the flag "count=TRUE", this shows now total reads, rather than proportion %
pclass2 <- ggbartax(obj=classtaxa, count=TRUE, facetNames="SEASON", topn=10) +
  xlab(NULL) +
  ylab("count reads") +
  scale_fill_manual(values=c(colorRampPalette(RColorBrewer::brewer.pal(12,"Set3"))(31))) +
  guides(fill= guide_legend(keywidth = 0.5, keyheight = 0.5))


pclass

pclass2


```


These following chunks might be best to compare metabarcoding vs traditional data but its OK to visualize below.
Way too many species difference between both sampling strategies and so the plots are messy  - after Kruskal, Wilcox and LDA, 264 taxa discriminate between wet and dry seasons.

```{r warning=FALSE}
library('coin')#for kruskal and wilcox test
# Since the effect size was calculated by randomly re-sampling, 
# the seed should be set for reproducibly results.
set.seed(1024)
deres <- diff_analysis(obj = ps_obj, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
deres
```

```{r MessyFigure_TaxoTree}
diffclade_p <- ggdiffclade(
  obj=deres, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=4, #taxonomy level from 1 to 8 kingdome:phylum:class:order:family:subfamily:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("goldenrod", "steelblue")
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_p
```


```{r BoxplotFIlteredFamily}
ps_obj %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^f__", f)) %>% ggdiffbox(colorlist=c("steelblue", "goldenrod"), notch = FALSE)

ps_obj %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^f__", f)) %>% ggeffectsize(colorlist=c("steelblue", "goldenrod"), notch = FALSE)

```


###### Please pay attention to the code below as it breaks down the code above, but to speed up the process so we don't have to convert form PS to MPSE to diff_anal. Instead it creates the object ps_obj_rareMPSE which I then used to filter by family and do aeither ggdifbox or ggeffectsize

ps_obj %>%
  as.MPSE() %>%
  mp_diff_analysis(.abundance = RareAbundance, .group = SEASON, action = 'get') -> ps_obj_rareMPSE

ps_obj_rareMPSE %>% 
  dplyr::filter(grepl("^f__", f)) %>% 
  ggdiffbox(colorlist = c ("steelblue", "goldenrod"), notch = FALSE)

ps_obj_rareMPSE %>% 
  dplyr::filter(grepl("^f__", f)) %>% 
  ggeffectsize(colorlist = c ("steelblue", "goldenrod"), notch = FALSE)

