---
title: "04_focal_order_seasons"
author: "Daniel"
date: "06/10/2022"
output: github_document
#editor_options: 
  #chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
rm(list=ls()) #I always start scritps this way to make sure I have a clean R environment
library('tidyr')
library('dplyr')
library('metacoder')
library('ggplot2')
library('agricolae')
library('vegan')
library('MicrobiotaProcess')
library('phyloseq')
library('ggtree')
library('coin')

obj <- readRDS('data/taxmap_object.rds') #loads the taxmap object created in script 02_metacoder_heat_trees

obj %>%  metacoder::filter_taxa(taxon_names %in% c("Lepidoptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> leps #we will create separate files for each order as this simplifies downstream analysis in microbiotaprocess - until I find a way to filter taxa on the mpse object directly.

obj %>%  metacoder::filter_taxa(taxon_names %in% c("Coleoptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> coleo
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Diptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> dips
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Hymenoptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> bees #i know its hymenoptera
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Hemiptera"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> hemi
obj %>%  metacoder::filter_taxa(taxon_names %in% c("Blattodea"),#here is to fliter the figure by groups
              subtaxa = TRUE) -> blats 

sample <- read.csv('data/location_ctrl.csv')
```

The script above gave me the taxmap object separated by order, however, it removes all taxonomic ranks up until "Order", so the script below is firstly, changing the taxmap object (e.g. leps) into a phyloseq object (as_phyloseq) and then using the microViz package to add the taxonomic ranks. 

There must be certainly a cleaner way how to do this without so many steps, but I then had to change the classification rank names again ("k__" instead of Kingdom) as this is more friendly with MicroBiotaProcess. Within the metabarcoding package community, there seems to have been several issues with the ranks, some have root, some don't, some have subfamily, some don't, so take note of the nomenclature you are using. I am using Kingdom;Phylum;Order;Class;Family;genus;species  (1 - 7).

If someone knows how to write a loop, this would be faster too. 
```{r CreatingOrderFiles message=FALSE, warning=FALSE, include=FALSE}
ps_leps <- metacoder::as_phyloseq(leps,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID") %>% microViz::tax_mutate(rank_0 = "Animalia",rank_01 = "Arthropoda", rank_02 = "Insecta", .before = 1) -> ps_leps 
colnames(tax_table(ps_leps)) <- c("k__", "p__", "o__", "c__", "f__", "g__", "s__")
  

ps_coleo <- metacoder::as_phyloseq(coleo,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID")  %>% microViz::tax_mutate(rank_0 = "Animalia",rank_01 = "Arthropoda", rank_02 = "Insecta", .before = 1) -> ps_coleo

colnames(tax_table(ps_coleo)) <- c("k__", "p__", "o__", "c__", "f__", "g__", "s__")

ps_dips <- metacoder::as_phyloseq(dips,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID")  %>% microViz::tax_mutate(rank_0 = "Animalia",rank_01 = "Arthropoda", rank_02 = "Insecta", .before = 1) -> ps_dips
colnames(tax_table(ps_dips)) <- c("k__", "p__", "o__", "c__", "f__", "g__", "s__")

ps_bees <- metacoder::as_phyloseq(bees,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID")  %>% microViz::tax_mutate(rank_0 = "Animalia",rank_01 = "Arthropoda", rank_02 = "Insecta", .before = 1) -> ps_bees
colnames(tax_table(ps_bees)) <- c("k__", "p__", "o__", "c__", "f__", "g__", "s__")

ps_blats <- metacoder::as_phyloseq(blats,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID")  %>% microViz::tax_mutate(rank_0 = "Animalia",rank_01 = "Arthropoda", rank_02 = "Insecta", .before = 1) -> ps_blats
colnames(tax_table(ps_blats)) <- c("k__", "p__", "o__", "c__", "f__", "g__", "s__")

ps_hemi <- metacoder::as_phyloseq(hemi,
                      otu_table = "tax_data",
                      otu_id_col = "bin_uri",
                      sample_data = sample,
                      sample_id_col = "sampleID")  %>% microViz::tax_mutate(rank_0 = "Animalia",rank_01 = "Arthropoda", rank_02 = "Insecta", .before = 1) -> ps_hemi
colnames(tax_table(ps_hemi)) <- c("k__", "p__", "o__", "c__", "f__", "g__", "s__")

```

Ok we now have datasets separated for each order (ForestGEO focal plus diptera - if we want more, we can adapt the script above to include more focal groups, or even to refine them)
But below we will create now some taxonomic trees showing differences between seasons (as in the MessyFigure chunk)

```{r Lepidoptera}
set.seed(1024)
deresleps <- diff_analysis(obj = ps_leps, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
deresleps

diffclade_leps <- ggdiffclade(
  obj=deresleps, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=5, #taxonomy level from 1 to 7 kingdom:phylum:class:order:family:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("orange", "darkcyan")
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_leps

ps_leps %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^f_", f)) %>% ggdiffbox(colorlist=c("darkcyan", "orange"), notch = FALSE)

es_leps <- ggeffectsize(obj=deresleps, 
                     lineheight=0.1,
                     linewidth=0.3, ytextsize = 0.3) + 
  scale_color_manual(values=c("orange", "darkcyan")) 

es_leps


```

```{r Coleoptera}
set.seed(1024)
derescoleo <- diff_analysis(obj = ps_coleo, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
derescoleo

diffclade_coleo <- ggdiffclade(
  obj=derescoleo, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=5, #taxonomy level from 1 to 7 kingdome:phylum:class:order:family:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("orange", "darkcyan")
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_coleo

ps_coleo %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^f_", f)) %>% ggdiffbox(colorlist=c("darkcyan", "orange"), notch = FALSE)

es_coleo <- ggeffectsize(obj=derescoleo, 
                     lineheight=0.1,
                     linewidth=0.3, ytextsize = 0.3) + 
  scale_color_manual(values=c("orange", "darkcyan")) 

es_coleo
```

```{r Diptera}
set.seed(1024)
deresdips <- diff_analysis(obj = ps_dips, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
deresdips

diffclade_dips <- ggdiffclade(
  obj=deresdips, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=5, #taxonomy level from 1 to 7 kingdome:phylum:class:order:family:subfamily:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("orange", "darkcyan")
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_dips

ps_dips %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^f_", f)) %>% ggdiffbox(colorlist=c("darkcyan", "orange"), notch = FALSE)

es_dips <- ggeffectsize(obj=deresdips, 
                     lineheight=0.1,
                     linewidth=0.3, ytextsize = 0.3) + 
  scale_color_manual(values=c("orange", "darkcyan")) 

es_dips

```



```{r Hemyptera}
set.seed(1024)
dereshemi <- diff_analysis(obj = ps_hemi, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
dereshemi

diffclade_hemi <- ggdiffclade(
  obj=dereshemi, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=5, #taxonomy level from 1 to 7 kingdome:phylum:class:order:family:subfamily:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("darkcyan", "orange" ) #Note that I had to change the order here, there are no hemiptera differences in DRY season so they did not come out in plot
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_hemi

ps_hemi %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^f_", f)) %>% ggdiffbox(colorlist=c("darkcyan", "orange"), notch = FALSE)

es_hemi <- ggeffectsize(obj=dereshemi, 
                     lineheight=0.1,
                     linewidth=0.3, ytextsize = 0.3) + 
  scale_color_manual(values=c( "darkcyan", "orange")) #same change due to wrong plotting 

es_hemi
```

There are problems with the Hymenoptera and Blattodea datasets. I still must figure it out. 

```{r Hymenoptera}
set.seed(1024)
deresbees <- diff_analysis(obj = ps_bees, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
deresbees

diffclade_bees <- ggdiffclade(
  obj=deresbees, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=5, #taxonomy level from 1 to 7 kingdome:phylum:class:order:family:subfamily:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("orange", "darkcyan")
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_bees

ps_bees %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^f_", f)) %>% ggdiffbox(colorlist=c("darkcyan", "orange"), notch = FALSE)

es_bees <- ggeffectsize(obj=deresbees, 
                     lineheight=0.1,
                     linewidth=0.3, ytextsize = 0.3) + 
  scale_color_manual(values=c("orange", "darkcyan")) 

es_bees

```

```{r Blattodea}
set.seed(1024)
deresblats <- diff_analysis(obj = ps_blats, classgroup = "SEASON",
                       mlfun = "lda",
                       filtermod = "pvalue",
                       firstcomfun = "kruskal_test",
                       firstalpha = 0.05,
                       strictmod = TRUE,
                       secondcomfun = "wilcox_test",
                       subclmin = 3,
                       subclwilc = TRUE,
                       secondalpha = 0.01,
                       lda=3,
                       action = "add")
deresblats

diffclade_blats <- ggdiffclade(
  obj=deresblats, 
  alpha=0.2, 
  linewd=0.15,
  skpointsize=0.2, 
  layout="radial",
  cladetext = 0.7,
  taxlevel=6, #taxonomy level from 1 to 8 kingdome:phylum:class:order:family:subfamily:genus:species
  removeUnkown=TRUE,
  reduce=TRUE # This argument is to remove the branch of unknown taxonomy.
) +
  scale_fill_manual(
    values=c("orange", "darkcyan")
  ) +
  guides(color = guide_legend(
    keywidth = 0.1, 
    keyheight = 0.6,
    order = 5,
    ncol=3)
  ) +
  theme(
    panel.background=element_rect(fill=NA),
    legend.position="right", 
    plot.margin=margin(0,0,0,0),
    legend.spacing.y=unit(0.02, "cm"), 
    legend.title=element_text(size=7),
    legend.text=element_text(size=6), 
    legend.box.spacing=unit(0.02,"cm")
  )
diffclade_blats

ps_blats %>% as.MPSE() %>% mp_rrarefy() %>% mp_diff_analysis(.abundance=RareAbundance, .group=SEASON, action='get') %>% dplyr::filter(grepl("^st_", f)) %>% ggdiffbox(colorlist=c("darkcyan", "orange"), notch = FALSE)

es_blats <- ggeffectsize(obj=deresblats, 
                     lineheight=0.1,
                     linewidth=0.3, ytextsize = 0.3) + 
  scale_color_manual(values=c("orange", "darkcyan")) 

es_blats
```
