---
title: "Rmarkdown"
author: "Daniel"
date: "16/09/2022"
output: github_document
---

I am now changing the markdown file to include my own script for BCI metabarcoding data.
This analysis relies hevily on tutorial for the metacoder package from the Grunwald Lab.
https://grunwaldlab.github.io/metacoder_documentation/example.html

```{r, message = FALSE}
#the small green box to insert chunks in different formats. 
#```{r} for R.
rm(list=ls()) #reoves all objects in workspace, start with clean environment
#load packages
library('tidyr')
library('dplyr')
library('metacoder')
```

First thing is to merge and clean the samples from the format downloaded from mbrave as right now its rather chaotic with column 1 sample, 2 BIN#, 3 #of seqs and 4 classificaiton.
```{r}
all_merged <- read.csv('data/light_trap_19_21.csv', header=T)
head(all_merged)
```

We need to group (and add) reads for the same BIN/classification in each sample

```{r}
all_merged %>%
  group_by(bin_uri, classification, sampleId) %>%
  summarise_all(sum) %>%
  data.frame() -> all_merged
write.csv(all_merged, 'data/dataset_merged.csv', row.names =FALSE)
```

This document now needs to be transposed (sample ID in columns, BIN_uri and classificaiton in rows). I did in excel, surely a way to do in R.

```{r}
merged_colnames <-read.csv('data/dataset_merged_samplecols.csv', header = T)
merged_colnames %>% #remove duplicated rows while adding reads to each 'sample' column
  group_by(bin_uri, classification) %>% #No more sampleID because they are the columns now
  summarise_all(sum) %>%
  data.frame() -> all_merged # so that newdf can further be used, if needed
write.csv(all_merged, 'data/merged_colnames_final.csv', row.names = F)
```

Now we have the final dataset, cleaned and formatted for use with Metacoder/MicroBiotaProcess